---
title: "Verb frequency and DOM among Heritage Speakers of Spanish "
subtitle: "⚔<br/>with xaringan"
author: "Esther Hur"
date: "2018/04/20"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---
#Verbal Lexical Frequency and DOM in Heritage Speakers of Spanish

```{r setup, include=FALSE}
library(tidyverse)
library(lme4)
library(lmerTest)
```
## **Who are Heritage Speakers?**

A subset of bilinguals who can be either simultaneous or sequential bilinguals “who learned a language other than English at home, but have been immersed in English since childhood, often exclusively schooled in it” (Wiley and Valdés, 2000, p. 80).

---
#Variability in HS Spanish

--Gender agreement (Montrul, Foote & Perpiñan, 2008)

--Preterite and imperfect use (Silva-Corvalán, 1994; Montrul, 2002; Montrul & Perpiñan, 2011) 

--Subjunctive mood morphology (Giancaspro, 2017; Montrul, 2009; Perez-Cortés, 2016)

--Differential Object Marking as expressed by ‘a’(Montrul 2004; Montrul and Bowles, 2009; Montrul & Sanchez-Walker, 2013; Montrul, Bhatt & Girju, 2015)

---

#What are the reasons?

--Incomplete Acquisition (Montrul, 2002; 2009)

--Attrition (Polinsky, 2006)

--Actication Approach (Putnam and Sanchez, 2013)

---
# What is the the Differential Object Marking (DOM)?


The overt marking of direct objects depending on the semantic and pragmatic features of the object. According to Bossong (1991), it is a phenomenon observed in at least 300 languages. It is conditioned by prominence: the higher in prominence a direct object, the more likely it is to be overtly case-marked.

---
### What about the Spanish DOM?

Marked Object | Unmarked object 
------------- | -------------------------------------------
[+animate]  |  [+animate]  |  [-animate]  |  [-animate]  
[+specific] |  [-specific] |  [+specific] | [-specific]
 
    (1)Cuid-           o                a      la niña

		  Take care-1.S.PRS (DOM) the girl
		  
		  'I'm taking care of the girl'
    (2)Cuid-               o      una niña

 		   Take care-1.S.PRS a girl
 		   
		  'I'm taking care of a girl'
    (3)Cuid-            o          el carro de mi hermano

	   	 Take care-1.S.PRS of my brother
	   	 
		  'I'm taking care of my brother's car'
---
# Incomplete Acquisition (Montrul 2002; 2009)

--How much input is sufficient enough for a complete acquisition?

--There is a need for a principled system of HS native language.

---

## RQ. Does lexical frequency of lexical activation play a role in the knowledge of DOM in animacy context among HS of Spanish in their productive knowledge?

---
# Participants

Advanced HS (Adv. HS) = 20

Intermediate HS (Int. HS) = 10

Spanish Dominant Controls (SDC) = 15

---
# Results

```{r, include=FALSE}

ept_df1 <- read_csv("../data/ept_raw.csv") %>% 
  mutate(., groupCon = recode(group, co = 0.5, hs = -0.5), #porque 0.5?
         freqCon = recode(frequency, f = 0.5, u = -0.5))

ept_df2 <- read_csv("../data/ept_additional_data.csv") %>% 
  mutate(., groupCon = recode(group, co = 0.5, hs = -0.5), 
         freqCon = recode(frequency, f = 0.5, u = -0.5))
         ept_df_temp <- rbind(ept_df1, ept_df2)


verbs <- c('f1' = 'cuidar', 
           'f2' = 'encontrar', 
           'f3' = 'tocar', 
           'f4' = 'aceptar', 
           'f5' = 'medir', 
           'f6' = 'parar', 
           'f7' = 'llevar', 
           'f8' = 'cambiar',
           'u1' = 'vigilar',
           'u2' = 'hallar', 
           'u3' = 'acariciar', 
           'u4' = 'acoger', 
           'u5' = 'abrigar', 
           'u6' = 'detener', 
           'u7' = 'trasladar', 
           'u8' = 'reemplazar')

ept_df_temp$verbs <- verbs[ept_df_temp$item]
ept_df <- read_csv("../data/prof_df.csv") %>% 
  left_join(ept_df_temp, ., by = 'participant')

# the new frequency counts from the Davis corpus
countfreq <- c('f1' = 7531,
               'f2' = 21725 , 
               'f3' = 3861  , 
               'f4' = 4098  , 
               'f5' = 886, 
               'f6' = 8174  , 
               'f7' = 107445 , 
               'f8' = 9871,
               'u1' = 1197 ,
               'u2' = 803, 
               'u3' = 427 , 
               'u4' = 2423, 
               'u5' = 123 , 
               'u6' = 5912, 
               'u7' = 4402, 
               'u8' = 4252)
ept_df$countfreq <- countfreq[ept_df$item]

ept_elog <- ept_df %>% 
  group_by(., participant, proficiency, countfreq, groupCon, response, item) %>% 
  summarize(., n = 8, 
            wDOM = sum(response), 
            woDOM = n - wDOM, 
            eLog = log((wDOM + 0.5) / (n - wDOM + 0.5)), 
            wts = 1 / (wDOM + 0.5) + 1 / (n - wDOM + 0.5),
            countLog = log(countfreq))
```
```{r table, echo=FALSE}

ept_df %>%
           group_by(., proficiency) %>%
           summarize(.,mean_response = mean(response), sd_score = sd(response),
                     mean_countfreq =mean(countfreq), sd_exp = sd(countfreq)) 
```
---
#### Figure 1 
```{r elog, echo=FALSE}
ept_elog2 <- ept_elog %>% filter(., item != "f2") %>%filter(., item != "f7") %>% filter(., item != "u3") %>% filter(., item != "u5")
```
```{r graph, fig.height=6, echo=FALSE}
ept_elog2 %>%
           ggplot(., aes(x = countLog, y = response, color = proficiency)) +
           geom_point()+
           geom_smooth(method =lm)+
           scale_color_brewer(palette = "Set1")
```
---
# Nested Model Comparison

```{r clean, include=FALSE}
ept_elog <- ept_df %>% 
  group_by(., participant, proficiency, countfreq, groupCon, response, item) %>% 
  summarize(., n = 8, 
            wDOM = sum(response), 
            woDOM = n - wDOM, 
            eLog = log((wDOM + 0.5) / (n - wDOM + 0.5)), 
            wts = 1 / (wDOM + 0.5) + 1 / (n - wDOM + 0.5),
            countLog = log(countfreq))
```  
```{r mod, include=FALSE}
mod_nul <- lm(eLog ~1, data = ept_elog2)
summary(mod_nul)

mod_cat <-lm(eLog ~ countLog, data = ept_elog2)
summary(mod_cat)

mod_con <-lm(eLog ~ countLog + proficiency, data = ept_elog2)
summary(mod_con)

mod_int <-lm(eLog ~ countLog * proficiency, data = ept_elog2)
summary(mod_int)
```
```{r anova, echo=FALSE}
anova(mod_nul, mod_cat, mod_con, mod_int)
```
---
